# StringHub

Starting from https://www.string-db.org/cgi/download

## Human

https://www.string-db.org/cgi/download?species_text=Homo+sapiens

This chunk is to be run manually, once

```{r eval=FALSE}
setwd("dev")

download.file("https://stringdb-static.org/download/protein.links.v11.5/9606.protein.links.v11.5.txt.gz", destfile = "9606.protein.links.v11.5.txt.gz")
download.file("https://stringdb-static.org/download/protein.links.detailed.v11.5/9606.protein.links.detailed.v11.5.txt.gz", destfile = "9606.protein.links.detailed.v11.5.txt.gz")
download.file("https://stringdb-static.org/download/protein.links.full.v11.5/9606.protein.links.full.v11.5.txt.gz", destfile = "9606.protein.links.full.v11.5.txt.gz")

download.file("https://stringdb-static.org/download/protein.info.v11.5/9606.protein.info.v11.5.txt.gz", destfile = "9606.protein.info.v11.5.txt.gz")
download.file("https://stringdb-static.org/download/protein.aliases.v11.5/9606.protein.aliases.v11.5.txt.gz", destfile = "9606.protein.aliases.v11.5.txt.gz")

```

```{r}
intdata_human <- vroom::vroom("9606.protein.links.v11.5.txt.gz")
intdata_human_detailed <- vroom::vroom("9606.protein.links.detailed.v11.5.txt.gz")
intdata_human_full <- vroom::vroom("9606.protein.links.full.v11.5.txt.gz")


colnames(intdata_human)
colnames(intdata_human_detailed)
colnames(intdata_human_full)

pryr::object_size(intdata_human)
pryr::object_size(intdata_human_detailed)
pryr::object_size(intdata_human_full)

# replace string identifiers

pinfo_human <- vroom::vroom("9606.protein.info.v11.5.txt.gz")
pinfo_human_detailed <- vroom::vroom("9606.protein.aliases.v11.5.txt.gz")

head(pinfo_human)
head(pinfo_human_detailed)

# working w a subset

ihfset <- intdata_human_full[1:2000, ]
head(ihfset)

hist(ihfset$combined_score, breaks = 100)

ihset <- intdata_human[1:2000, ]
head(ihset)

ihset$p1_genename <- pinfo_human$preferred_name[match(ihset$protein1, pinfo_human$`#string_protein_id`)]
ihset$p2_genename <- pinfo_human$preferred_name[match(ihset$protein2, pinfo_human$`#string_protein_id`)]

# thought: adding entrez ids as well? for better interoperability?
  
```

General info

https://stringdb-static.org/download/species.v11.5.txt

### Via STRINGdb

```{r}
library(STRINGdb)

# needs to specify explicitly, could be autocreated
dir.create("sdb_human")
string_db <- STRINGdb$new(version="11", species=9606,
                          score_threshold=200, input_directory="sdb_human")

STRINGdb$methods()
STRINGdb$help("get_graph")

sdbg <- string_db$get_graph()

```

### From the manually downloaded files

Becomes a little more transparent what really goes on

```{r}
PPI <- read.table("sdb_human/9606.protein.links.v11.0.txt.gz", sep = " ",
                  header=TRUE, stringsAsFactors=FALSE, fill = TRUE)
pryr::object_size(sdbg)
pryr::object_size(PPI)
```

### Focusing on a subset of the graph

Using the example in STRINGdb to showcase what I would like to do often...

```{r}
data(diff_exp_example1)
head(diff_exp_example1)
example1_mapped <- string_db$map(diff_exp_example1, "gene", 
                                 removeUnmappedRows = TRUE)

hits <- example1_mapped$STRING_id[1:200]
hits

string_db$plot_network(hits)

score_threshold <- 200

PPIselected <- PPI[PPI$combined_score >= score_threshold,]
dim(PPI)
dim(PPIselected)

library("igraph")

sdbg

myg <- igraph::graph.data.frame(d = PPIselected, directed = FALSE)
myg
sdbg

subg <- igraph::induced.subgraph(sdbg, which(V(sdbg)$name %in% hits))
subg

# 784 is 392*2, am I using the subgraph function wrong?

# why 199 vs 200?!

intersect(V(sdbg)$name, hits)
```

Starting to have fun with interactive networks...

```{r}
visNetwork::visIgraph(subg)

V(subg)
str(V(subg))

# putting on top some more interesting/readable info
node_info <- vroom::vroom("9606.protein.info.v11.5.txt.gz")
node_info$preferred_name[match(V(subg)$name, node_info$`#string_protein_id`)]

V(subg)$name <- 
  node_info$preferred_name[match(V(subg)$name, node_info$`#string_protein_id`)]

visNetwork::visIgraph(subg)

library("magrittr")
library("visNetwork")
visNetwork::visIgraph(subg) %>%
  visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)

# re-sort the nodes alphabetically? makes sense
rank_gs <- rank(V(subg)$name)
subg <- permute.vertices(subg, rank_gs)
visNetwork::visIgraph(subg) %>%
  visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)
```

With a smaller set, to check in detail:

```{r}
hits <- example1_mapped$STRING_id[1:100]
hits
string_db$plot_network( hits )
subg <- igraph::induced.subgraph(sdbg, which(V(sdbg)$name %in% hits))
subg

V(subg)$name <- node_info$preferred_name[match(V(subg)$name, node_info$`#string_protein_id`)]

# re-sort the nodes alphabetically? makes sense
rank_gs <- rank(V(subg)$name)
subg <- permute.vertices(subg, rank_gs)

visNetwork::visIgraph(subg) %>%
  visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)

# double edges amount?!
subg <- simplify(subg)
```

Now: putting on top the info from the expression values and their changes

```{r}
hits
# different colors for the nodes in function of their logFC
set.seed(42)
fcs_genes <- rnorm(mean = 0, sd = 3, n = length(hits))
fcs_genes

mypal <- rev(scales::alpha(
  colorRampPalette(RColorBrewer::brewer.pal(name = "RdYlBu", 11))(50), 0.4
))

V(subg)$color <- GeneTonic::map2color(fcs_genes, mypal, limits = c(-4, 4))

visNetwork::visIgraph(subg) %>%
  visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)

# title for tooltips
V(subg)$title <- NA

V(subg)$title <- paste0(
  "<h4>", V(subg)$name, "</h4><br>",
  "logFC = ", format(round(fcs_genes, 2), nsmall = 2)
)

visNetwork::visIgraph(subg) %>%
  visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)
```

Coloring "different things" - a la payload, using the halo?

```{r}
subg <- igraph::induced.subgraph(sdbg, which(V(sdbg)$name %in% hits))
subg
subg <- igraph::simplify(subg)
V(subg)$name <- node_info$preferred_name[match(V(subg)$name, node_info$`#string_protein_id`)]
# re-sort the nodes alphabetically? makes sense
rank_gs <- rank(V(subg)$name)
subg <- permute.vertices(subg, rank_gs)

V(subg)$borderWidth  <- 5
V(subg)$color.border <- GeneTonic::map2color(fcs_genes, mypal, limits = c(-4, 4))
visNetwork::visIgraph(subg) %>%
  visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)

# title for tooltips
V(subg)$title <- NA
V(subg)$title <- paste0(
  "<h4>", V(subg)$name, "</h4><br>",
  "logFC = ", format(round(fcs_genes, 2), nsmall = 2)
)

visNetwork::visIgraph(subg) %>%
  visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)

```


