---
title: "NetworkHub: working with network data, seamlessly"
author: 
- name: Annekathrin Ludt
  affiliation: 
  - &id1 Institute of Medical Biostatistics, Epidemiology and Informatics (IMBEI), Mainz
  email: anneludt@uni-mainz.de
- name: Federico Marini
  affiliation: 
  - *id1
  email: marinif@uni-mainz.de
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
    code_download: true
    global_numbering: true
    number_sections: false
editor_options: 
  chunk_output_type: console
bibliography: NetworkHub.bib
link-citations: true
date: "`r BiocStyle::doc_date()`"
package: "`r BiocStyle::pkg_ver('NetworkHub')`"
vignette: >
  %\VignetteIndexEntry{networkhub_string}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center",
  fig.show = "asis",
  eval = FALSE,
  fig.width = 10,
  fig.height = 7,
  tidy = FALSE,
  message = FALSE,
  warning = FALSE,
  size = "small",
  echo = TRUE,
  results = "markup"
)

options(
  replace.assign = TRUE, 
  width = 100
)
```

# What is NetworkHub?

and why...

# Setting up NetworkHub

install
setup
explain cache

```{r}
initialize_NetworkHub(nh_cachedir = "NetworkHub")
```


# What formats are available?

At large
..
and what we use here

What are our "output" formats?


# Using NetworkHub

## STRING

This is to obtain a relatively fully fledged snapshot of what is out there for some widely used core organisms.

```{r}
# some form of mass fetching ----------------------------------------------

info_stringdb_species() |> 
  dplyr::filter(STRING_type == "core")

main_core_organisms <- c(
  "Homo sapiens",
  "Mus musculus",
  "Arabidopsis thaliana",
  "Saccharomyces cerevisiae",
  "Danio rerio",
  "Drosophila melanogaster",
  "Caenorhabditis elegans",
  "Pseudomonas aeruginosa PAO1",
  "Escherichia coli str. K-12 substr. MG1655"
)

main_core_organisms %in% info_stringdb_species()$official_name_NCBI

lapply(main_core_organisms, function(species) {
  get_networkdata_STRINGDB(species = species,
                           version = "11.5")
})  

lapply(main_core_organisms, function(species) {
  get_accessoryinfo_STRINGDB(species = species,
                             version = "11.5")
})  

```

Bringing the content into a graph

```{r}
sdb_human <- get_networkdata_STRINGDB("Homo sapiens",
                                      version = "11.5")
dim(sdb_human)
```

Load in for example the results from differential expression

```{r}
# res object
data(res_de_macrophage, package = "GeneTonic")
res_de <- res_macrophage_IFNg_vs_naive
first_degenes <- na.omit(GeneTonic::deseqresult2df(res_de)[1:150, "id"])

library("macrophage")
library("DESeq2")
library("org.Hs.eg.db")
library("AnnotationDbi")

# dds object
data("gse", package = "macrophage")
dds_macrophage <- DESeqDataSet(gse, design = ~ line + condition)
rownames(dds_macrophage) <- substr(rownames(dds_macrophage), 1, 15)
dds_macrophage <- estimateSizeFactors(dds_macrophage)

se_macrophage <- 
  DeeDee::DeeDeeExperiment(
    dds_macrophage, 
    de_results = list(
      macrores = res_de
    )
  )

se_macrophage
```

Going from the DE genes to the proteins...

```{r}
aux_info_string <- get_accessoryinfo_STRINGDB(species = "Homo sapiens",
                             version = "11.5")

anno_df_string <- create_annotation_from_STRINGDBaccessory(aux_info_string)

head(anno_df_string)

first_proteins <- anno_df_string$protein_id[match(first_degenes, anno_df_string$ensembl_id)] |> 
  na.omit()
```

Getting a subnetwork for the DE proteins...

```{r}
de_proteins_subgraph <- build_graph_STRINGDB(graph_data = sdb_human, 
                                             output_format = "igraph",
                                             min_score_threshold = 200,
                                             subset_nodes = first_proteins)

plot(de_proteins_subgraph)

# re-sort the nodes alphabetically? makes sense
rank_gs <- rank(V(de_proteins_subgraph)$name)
de_proteins_subgraph <- permute.vertices(de_proteins_subgraph, rank_gs)
visNetwork::visIgraph(de_proteins_subgraph) %>%
  visNetwork::visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)

V(de_proteins_subgraph)
str(V(de_proteins_subgraph))
```

Starting to have fun with interactive networks...  
Makes sense to display the gene names

```{r}

de_proteins_subgraph

de_graph <- de_proteins_subgraph

# renaming the nodes, using the gene symbols
V(de_graph)$name <- 
  anno_df_string$ensembl_id[match(V(de_graph)$name, anno_df_string$protein_id)]

de_graph <- igraph::induced.subgraph(de_graph, which(!is.na(V(de_graph)$name)))

# re-sort the nodes alphabetically? makes sense
rank_gs <- rank(V(de_graph)$name)
de_graph <- permute.vertices(de_graph, rank_gs)
visNetwork::visIgraph(de_graph) %>%
  visNetwork::visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)
```

Now: putting on top the info from the expression values and their changes

```{r}
# different colors for the nodes in function of their logFC
# set.seed(42)
# fcs_genes <- rnorm(mean = 0, sd = 3, n = length(V(de_graph)))

str(V(de_graph))

intersect(V(de_graph)$name, first_degenes)

V(de_graph)$log2FC

V(de_graph)$log2FC <- res_de$log2FoldChange[match(V(de_graph)$name, rownames(res_de))]

mypal <- rev(scales::alpha(
  colorRampPalette(RColorBrewer::brewer.pal(name = "RdBu", 11))(50), 0.4
))

max_de <- max(abs(range(V(de_graph)$log2FC)))

V(de_graph)$color <- GeneTonic::map2color(V(de_graph)$log2FC, mypal, limits = c(-max_de, max_de))

gene_ids_backup <- V(de_graph)$name
gene_names <- anno_df_string$gene_symbol[match(gene_ids_backup, anno_df_string$ensembl_id)]
new_names <- gene_names
new_names[is.na(new_names)] <- gene_ids_backup[is.na(new_names)]
V(de_graph)$name <- new_names

rank_gs <- rank(V(de_graph)$name)
de_graph <- permute.vertices(de_graph, rank_gs)

# title for tooltips
V(de_graph)$title <- NA

V(de_graph)$title <- paste0(
  "<h4>", V(de_graph)$name, "</h4><br>",
  "logFC = ", format(round(V(de_graph)$log2FC, 2), nsmall = 2)
)

visNetwork::visIgraph(de_graph) |> 
  visNetwork::visEdges(color = list(color = "#88888888")) |> 
  visNetwork::visOptions(highlightNearest = list(enabled = TRUE,
                                                 degree = 1,
                                                 hover = TRUE),
                         nodesIdSelection = TRUE)
```

Coloring "different things" - a la payload, using the halo?

```{r}
subg <- igraph::induced.subgraph(sdbg, which(V(sdbg)$name %in% hits))
subg
subg <- igraph::simplify(subg)
V(subg)$name <- node_info$preferred_name[match(V(subg)$name, node_info$`#string_protein_id`)]
# re-sort the nodes alphabetically? makes sense
rank_gs <- rank(V(subg)$name)
subg <- permute.vertices(subg, rank_gs)

V(subg)$borderWidth  <- 5
V(subg)$color.border <- GeneTonic::map2color(fcs_genes, mypal, limits = c(-4, 4))
visNetwork::visIgraph(subg) %>%
  visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)

# title for tooltips
V(subg)$title <- NA
V(subg)$title <- paste0(
  "<h4>", V(subg)$name, "</h4><br>",
  "logFC = ", format(round(fcs_genes, 2), nsmall = 2)
)

visNetwork::visIgraph(subg) %>%
  visOptions(highlightNearest = list(enabled = TRUE,
                                     degree = 1,
                                     hover = TRUE),
             nodesIdSelection = TRUE)

```


## BioGrid


## IntAct


## HIPPIE


## From KEGG?



# A note on the identifiers

protein ids, 
ensembl,
symbol

# NetworkHub in existing workflows

## Projecting DE results onto a network


## Usage with other software packages requiring 


# NetworkHub on the Bioc HUBS


# Session Info {-}

```{r}
sessionInfo()
```


<hr>

# References

